Enforced Invariant Confluence

### 2018-05-01 doc/paper/sections/abstract.tex
Couple more tweaks.

\begin{abstract}
Strongly consistent systems are easy to reason about but face fundamental
limitations in availability and performance. Weakly consistent systems can be
implemented with very high performance but place a burden on the application
developer to reason about complex interleavings of execution.
\Invariantconfluence{} provides a formal framework for understanding when we
can get the best of both worlds. An \invariantconfluent{} object can be
efficiently replicated with no coordination needed to preserve its invariants.
However, actually determining whether or not an object is \invariantconfluent{}
is challenging (undecidable). In this paper, we establish conditions under
which a commonly used sufficient condition for \invariantconfluence{} is both
necessary and sufficient and use this condition to design (a) a general-purpose
interactive \invariantconfluence{} decision procedure and (b) a novel
sufficient condition which can be checked automatically. We then take a step
beyond \invariantconfluence{} and introduce a generalization of
\invariantconfluence{} called segmented \invariantconfluence{}, which allows us
to replicate non-\invariantconfluent{} objects with a small and tunable amount
of coordination. We implemented our theoretical findings in a prototype called
Lucy and found that our decision procedures efficiently handle workloads
including foreign key constraints, rollups, escrow transactions, etc.  We also
found that segmented \invariantconfluent{} replication can deliver up to an
order of magnitude more throughput than linearizable replication for low
contention workloads and comparable throughput for medium to high contention
workloads.
\end{abstract}

>>>>

\begin{abstract}
Strongly consistent systems are easy to reason about but face fundamental
limitations in availability and performance. Weakly consistent systems can be
implemented with very high performance but place a burden on the application
developer to reason about complex interleavings of execution.
\Invariantconfluence{} provides a formal framework for understanding when we
can get the best of both worlds. An \invariantconfluent{} object can be
efficiently replicated with no coordination needed to preserve its invariants.
However, actually determining whether or not an object is \invariantconfluent{}
is challenging (undecidable). In this paper, we establish conditions under
which a commonly used sufficient condition for \invariantconfluence{} is both
necessary and sufficient and use this condition to design (a) a general-purpose
interactive \invariantconfluence{} decision procedure and (b) a novel
sufficient condition which can be checked automatically. We then take a step
beyond \invariantconfluence{} and introduce a generalization of
\invariantconfluence{} called segmented \invariantconfluence{}, which allows us
to replicate non-\invariantconfluent{} objects with a small amount of
coordination. We implemented our theoretical findings in a prototype called
Lucy and found that our decision procedures efficiently handle workloads
including foreign key constraints, rollups, escrow transactions, etc.  We also
found that segmented \invariantconfluent{} replication can deliver up to an
order of magnitude more throughput than linearizable replication for low
contention workloads and comparable throughput for medium to high contention
workloads.
\end{abstract}

### 2018-05-01 doc/paper/sections/evaluation.tex
Added eval table.
Couple more tweaks.

\subsection{Decision Procedures}
We now evaluate the practicality and efficiency of our decision procedure
prototypes. We begin by demonstrating the decision procedure on a handful of
simple, yet practical examples. We then discuss how our tool can be used to
analyze the TPC-C benchmark.

@@

  % \item
    Two of the twelve invariants involve generating \textbf{sequential and
    unique identifiers}. Consider, for example, invariant 2 which dictates that
    a district's next order id \ttt{D\_NEXT\_O\_ID} is equal to the maximum
    order id \ttt{O\_ID} of orders within the district. The New Order
    transaction places an order with \ttt{O\_ID} equal to the current value of
    \ttt{D\_NEXT\_O\_ID} and then increments \ttt{D\_NEXT\_O\_ID}.
    We model this workload with an integer for \ttt{D\_NEXT\_O\_ID} and a map
    for \ttt{O\_ID} that maps order ids to order.
    %
    % We model this workload with three objects. \ttt{D\_NEXT\_O\_ID} is modeled
    % as an integer merged by $\max$. The order table is modelled as two objects.
    % The first is a set of all order ids used, merged by union. The second is a
    % map which maps order id to order. The map's merge function performs an
    % id-by-id merge of orders. If two different orders have the same id, the
    % merge function replaces them with a distinguished value to indicate that
    % the primary key constraint has been violated.
    %
    We applied Lucy to this workload and in less than a second, it produced a
    counterexample that---when labelled as reachable---confirming that the
    workload is \emph{not} \invariantconfluent{}~\cite{bailis2014coordination}.
% \end{enumerate}

@@

  Second, throughput does not decrease linearly with the amount of
  coordination. Even infrequent coordination can drastically decrease
  throughput. Increasing the fraction of decrements from 0\% to 1\% decreases
  throughput by a factor of 2. Increasing again to 3\%, the throughput
  decreases by another factor of 2. With 90\% coordination-free transactions
  (i.e.\ 10\% decrements), we achieve only 10\% of the throughput of eventual
  consistency.

@@

  These results echo the results of \benchref{VaryWithdraws}. For
  low-coordination workloads, segmented \invariantconfluent{} replication can
  offer almost an order of magnitude better scalability of linearizable
  replication, but coordination decreases scalability superlinearly. Even
  infrequent coordination can drastically reduce the scalability of segmented
  \invariantconfluent{} replication with segmented \invariantconfluent{}
  replication ultimately scaling worse than linearizable replication for
  high-coordination workloads.
\end{benchmark}

>>>>

\subsection{Decision Procedures}
We now evaluate the practicality and efficiency of our decision procedure
prototypes. We begin by demonstrating the decision procedure on a handful of
simple, yet practical examples. We then discuss how our tool can be used to
analyze the TPC-C benchmark. A summary of these results is given in
\tabref{EvalTable}.

{\input{figures/eval_table.tex}}

@@

  % \item
    Two of the twelve invariants involve generating \textbf{sequential and
    unique identifiers}. Consider, for example, invariant 2 which dictates that
    a district's next order id \ttt{D\_NEXT\_O\_ID} is equal to the maximum
    order id \ttt{O\_ID} of orders within the district. The New Order
    transaction places an order with \ttt{O\_ID} equal to the current value of
    \ttt{D\_NEXT\_O\_ID} and then increments \ttt{D\_NEXT\_O\_ID}.
    We model this workload with an integer for \ttt{D\_NEXT\_O\_ID} and a map
    for \ttt{O\_ID} that maps order ids to order.
    %
    % We model this workload with three objects. \ttt{D\_NEXT\_O\_ID} is modeled
    % as an integer merged by $\max$. The order table is modelled as two objects.
    % The first is a set of all order ids used, merged by union. The second is a
    % map which maps order id to order. The map's merge function performs an
    % id-by-id merge of orders. If two different orders have the same id, the
    % merge function replaces them with a distinguished value to indicate that
    % the primary key constraint has been violated.
    %
    We applied Lucy to this workload and in less than a second, it produced a
    counterexample that---when labelled as reachable---confirms that the
    workload is \emph{not} \invariantconfluent{}~\cite{bailis2014coordination}.
% \end{enumerate}

@@

  Second, throughput does not decrease linearly with the amount of
  coordination. Even infrequent coordination can drastically decrease
  throughput. Increasing the fraction of decrements from 0\% to 1\% decreases
  throughput by a factor of 2. Increasing again to 3\%, the throughput
  decreases by another factor of 2. With 90\% coordination-free transactions
  (i.e.\ 10\% decrements), we achieve only 10\% of the throughput of eventually
  consistent replication.

@@

  These results echo the results of \benchref{VaryWithdraws}. For
  low-coordination workloads, segmented \invariantconfluent{} replication can
  offer almost an order of magnitude better scalability compared to
  linearizable replication, but coordination decreases scalability
  superlinearly. Even infrequent coordination can drastically reduce the
  scalability of segmented \invariantconfluent{} replication with segmented
  \invariantconfluent{} replication ultimately scaling worse than linearizable
  replication for high-coordination workloads.
\end{benchmark}

### 2018-05-01 doc/paper/sections/interactive_decision_procedure.tex
Moved figure around.
Couple more tweaks.

\section{Interactive Decision Procedure}\seclabel{InteractiveDecisionProcedure}
\newcommand{\Helper}{\textsf{Helper}}
\newcommand{\IsIclosed}{\textsf{IsIclosed}}
\newcommand{\IsInvConfluent}{\textsf{IsInvConfluent}}

\newcommand{\algocomment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
\begin{algorithm}[t]
  \caption{Interactive invariantconfluence{} decision procedure}%
  \algolabel{InteractiveDecisionProcedure}
  \begin{algorithmic}
    \algocomment{Return if $O$ is \sTIconfluent{}.}
    \Function{IsInvConfluent}{$O$, $s_0$, $T$, $I$}
      \State
        \Return $I(s_0)$ and
        \Call{Helper}{$O$, $s_0$, $T$, $I$, $\set{s_0}$, $\emptyset$}
    \EndFunction

@@

\thmref{IclosureEquivalentIconfluence} tells us that if all invariant
satisfying points are reachable, then \invariantclosure{} and
invariantconfluence{} are equivalent. In this section, we present the
interactive invariantconfluence{} decision procedure shown in
\algoref{InteractiveDecisionProcedure} which takes advantage of this result.

\subsection{The Decision Procedure}
A user provides \algoref{InteractiveDecisionProcedure} with an object $O = (S,
\join)$, a start state $s_0$, a set of transactions $T$, and an invariant
$I$\footnote{%
  Formally, $I$ is a (possibly infinite) set of states (e.g. $\setst{(x, y) \in
  \ints \times \ints}{xy \leq 0}$), but in practice, our implementation of
  \IsInvConfluent{} takes in $I$ as a finite logical formula (e.g. $xy \leq
  0$).
}. The user then interacts with
\algoref{InteractiveDecisionProcedure} to iteratively eliminate unreachable
states from the invariant. Meanwhile, the algorithm leverages an
\invariantclosure{} decision procedure to either (a) conclude that $O$ is or is
not \sTIconfluent{} or (b) provide counterexamples to the user to help them
eliminate unreachable states. After all unreachable states have been eliminated
from the invariant, \thmref{IclosureEquivalentIconfluence} allows us to reduce
the problem of invariantconfluence{} directly to the problem of
\invariantclosure{}, and the algorithm terminates.
%
We now describe \algoref{InteractiveDecisionProcedure} in detail. An example of
how to use \algoref{InteractiveDecisionProcedure} on \exampleref{Z2} is given
in \figref{DecisionProcedureExample}.

@@

\IsInvConfluent{} first checks that $s_0$ satisfies the invariant. $s_0$ is
reachable, so if it does not satisfy the invariant, then $O$ is not
\sTIconfluent{} and \IsInvConfluent{} returns false. Otherwise,
\IsInvConfluent{} calls a helper function \Helper{} which---in addition to $O$,
$s_0$, $T$, and $I$---takes as arguments a set $R$ of \sTIreachable{} states
and a set $NR$ of \sTIunreachable{} states. Like \IsInvConfluent, \Helper$(O,
s_0, T, I, R, NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and
$NR$ are correct).  As \algoref{InteractiveDecisionProcedure} executes, $NR$ is
iteratively increased, which removes unreachable states from $I$ until $I
\subseteq \setst{s \in S}{\sTIreachablepredicate{s}}$.

>>>>

\section{Interactive Decision Procedure}\seclabel{InteractiveDecisionProcedure}
\newcommand{\Helper}{\textsf{Helper}}
\newcommand{\IsIclosed}{\textsc{IsIclosed}}
\newcommand{\IsInvConfluent}{\textsc{IsInvConfluent}}

\newcommand{\algocomment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
\begin{algorithm}[t]
  \caption{Interactive \invariantconfluence{} decision procedure}%
  \algolabel{InteractiveDecisionProcedure}
  \begin{algorithmic}
    \algocomment{Return if $O$ is \sTIconfluent{}.}
    \Function{IsInvConfluent}{$O$, $s_0$, $T$, $I$}
      \State
        \Return $I(s_0)$ and
        \Call{Helper}{$O$, $s_0$, $T$, $I$, $\set{s_0}$, $\emptyset$}
    \EndFunction

@@

\thmref{IclosureEquivalentIconfluence} tells us that if all invariant
satisfying points are reachable, then \invariantclosure{} and
\invariantconfluence{} are equivalent. In this section, we present the
interactive \invariantconfluence{} decision procedure shown in
\algoref{InteractiveDecisionProcedure} which takes advantage of this result.

\subsection{The Decision Procedure}
A user provides \algoref{InteractiveDecisionProcedure} with an object $O = (S,
\join)$, a start state $s_0$, a set of transactions $T$, and an invariant
$I$\footnote{%
  Formally, $I$ is a (possibly infinite) set of states (e.g. $\setst{(x, y) \in
  \ints \times \ints}{xy \leq 0}$), but in practice, our implementation of
  \IsInvConfluent{} takes in $I$ as a finite logical formula (e.g. $xy \leq
  0$).
}. The user then interacts with
\algoref{InteractiveDecisionProcedure} to iteratively eliminate unreachable
states from the invariant. Meanwhile, the algorithm leverages an
\invariantclosure{} decision procedure to either (a) conclude that $O$ is or is
not \sTIconfluent{} or (b) provide counterexamples to the user to help them
eliminate unreachable states. After all unreachable states have been eliminated
from the invariant, \thmref{IclosureEquivalentIconfluence} allows us to reduce
the problem of \invariantconfluence{} directly to the problem of
\invariantclosure{}, and the algorithm terminates.
%
We now describe \algoref{InteractiveDecisionProcedure} in detail. An example of
how to use \algoref{InteractiveDecisionProcedure} on \exampleref{Z2} is given
in \figref{DecisionProcedureExample}.

@@

\IsInvConfluent{} first checks that $s_0$ satisfies the invariant. $s_0$ is
reachable, so if it does not satisfy the invariant, then $O$ is not
\sTIconfluent{} and \IsInvConfluent{} returns false. Otherwise,
\IsInvConfluent{} calls a helper function \Helper{} which---in addition to $O$,
$s_0$, $T$, and $I$---takes as arguments a set $R$ of \sTIreachable{} states
and a set $NR$ of \sTIunreachable{} states. Like \IsInvConfluent, \Helper$(O,
s_0, T, I, R, NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and
$NR$ are correct).  As \algoref{InteractiveDecisionProcedure} executes, $NR$ is
iteratively increased, which removes unreachable states from $I$ until $I$ is a
subset of $\setst{s \in S}{\sTIreachablepredicate{s}}$.

### 2018-05-01 doc/paper/sections/segmented_invariant_confluence.tex
Couple more tweaks.

\section{Segmented Invariant Confluence}\seclabel{SegmentedInvariantConfluence}
\newcommand{\IsIclosed}{\textsf{IsIclosed}}

@@

\begin{example}\examplelabel{SegmentedZ2}
  Consider again the object $O = (\ints \times \ints, \join)$, transactions $T
  = \set{t_{x+1}, t_{y-1}}$, and invariant $I = \setst{(x, y)}{xy \leq 0}$ from
  \exampleref{Z2}, but now let the start state $s_0$ be $(-42, 42)$. $O$ is
  \emph{not} \sTISconfluent{} because the points $(0, 42)$ and $(42, 0)$ are
  reachable, and merging these points yields $(42, 42)$ which violates the
  invariant. However, $O$ is \sTISconfluent{} for $\Sigma = (I_1, T_1)$, $(I_2,
  T_2)$, $(I_3, T_3)$, $(I_4, T_4)$ where
  \begin{align*}
    I_1 &= \setst{(x, y)}{x < 0, y > 0}       &T_1 &= \set{t_{x+1}, t_{y-1}} \\
    I_2 &= \setst{(x, y)}{x \geq 0, y \leq 0} &T_2 &= \set{t_{x+1}, t_{y-1}} \\
    I_3 &= \setst{(x, y)}{x = 0}              &T_3 &= \set{t_{y-1}} \\
    I_4 &= \setst{(x, y)}{y = 0}              &T_4 &= \set{t_{x+1}}
  \end{align*}
  $\Sigma$ is illustrated in \figref{SegmentedZ2}. Clearly, $s_0$ satisfies the
  invariant, and $I_1, I_2, I_3, I_4$ cover $I$. Moreover, for every $(I_i,
  T_i) \in \Sigma$, we see that $O$ is \iclosed{I_i}, so $O$ is
  \sticonfluent{s}{T_i}{I_I} for every $s \in I_i$. Thus, $O$ is
  \sTISconfluent{}.
\end{example}

@@

If $t \notin T_i$, then $p_i$ initiates a round of global coordination to
execute $t$. During global coordination, every server temporarily stops
processing and transitions to state $s = s_1 \join \ldots \join s_n$, the join
of every server's state. Then, every server speculatively executes $t$
transitioning to state $t(s)$. If $t(s)$ violates the invariant (i.e.  $\lnot
I(t(s))$), then every server aborts $t$ and reverts to state $s$.  Then, $p_i$
replies to the client. If $t(s)$ satisfies the invariant (i.e. $I(t(s))$), then
every server commits $t$ and remains in state $t(s)$. Every server then chooses
the first segment $(I_i, T_i) \in \Sigma$ such that $I_i(t(s))$ to be the new
active segment. Note that such a segment is guaranteed to exist because the
segment invariants cover $I$. Moreover, $\Sigma$ is ordered, so every server
will deterministically pick the same active segment. In fact, an invariant of
the system model is that at any given point of normal processing, every server
has the same active segment.

If $t \in T_i$, then $p_i$ executes $t$ immediately and without coordination.
If $t(s_i)$ satisfies the \emph{active} invariant (i.e. $I_i(t(s_i))$), then
$p_i$ commits $t$, stays in state $t(s_i)$, and replies to the client. If
$t(s_i)$ violates the invariant (i.e. $\lnot I(t(s_i))$), then $p_i$ aborts
$t$, reverts to state $s_i$, and replies to the client. If $t(s_i)$ satisfies
the invariant but violates the active invariant (i.e. $I(t(s_i))$ but $\lnot
I_i(t(s_i))$), then $p_i$ reverts to state $s_i$ and initiates a round of
global coordination to execute $t$, as described in the previous paragraph.
Transaction execution is summarized in \algoref{TxnExecution}.

@@

\subsection{Interactive Decision Procedure}
In order for us to determine whether or not an object $O$ is \sTISconfluent{},
we have to determine whether or not $O$ is \invariantconfluent{} within each
segment $(I_i, T_i) \in \Sigma$. That is, we have determine if $O$ is
\sticonfluent{s}{T_i}{I_i} confluent for every state $s \in I_i$. Ideally, we
could leverage \algoref{InteractiveDecisionProcedure}, invoking it once per
segment. Unfortunately, \algoref{InteractiveDecisionProcedure} can only be used
to determine if $O$ is \sticonfluent{s}{T_i}{I_i} for a \emph{particular} state
$s \in I_i$, not for \emph{every} state $s \in I_i$. Thus, we would have to
invoke \algoref{InteractiveDecisionProcedure} $|I_i|$ times for every segment
$(I_i, T_I)$, which is clearly infeasible given that $I_i$ can be large or even
infinite.

@@

Using \thmref{ExtendedIclosureEquivalentIconfluence}, we develop
\algoref{ArbitraryStartInteractiveDecisionProcedure}: a natural generalization
of \algoref{InteractiveDecisionProcedure}.
\algoref{InteractiveDecisionProcedure} iteratively refines the set of
\emph{reachable} states whereas
\algoref{ArbitraryStartInteractiveDecisionProcedure} iteratively refines the
set of \emph{coreachable} states, but otherwise, the core of the two algorithms
is the same.\footnote{%
  Another small difference is that \IsIclosed{} behaves differently in
  \algoref{InteractiveDecisionProcedure} and
  \algoref{ArbitraryStartInteractiveDecisionProcedure}. In
  \algoref{ArbitraryStartInteractiveDecisionProcedure}, \IsIclosed{} returns a
  triple (closed, $s_1$, $s_2$) where if closed is false, then $s_1, s_2 \in I$
  are two states not in $NR$ such that $I(s_1)$ and $I(s_2)$ but $\lnot I(s_1
  \join s_2)$. If no such states exist, then closed is true, and $s_1$ and
  $s_2$ are null.
}
Now, a segmented \invariantconfluence{} decision procedure, can simply invoke
\algoref{ArbitraryStartInteractiveDecisionProcedure} once on each segment.

@@

Third, we can integrate a couple of optimizations into our system model to
further reduce the amount of coordination required. First, if a server with
state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t \in I_i$
to execute, and $t(s_i)$ violates the active invariant but not the global
invariant, instead of initiating a round of global coordination, $p_i$ can
simply buffer $t$ for re-execution at a later time. While this increases the
latency required to execute $t$, it's possible that after other transactions
are executed, re-executing $t$ may lead to a state that either satisfies the
active invariant or violates the global invariant. In either case, a round of
global coordination is avoided. Similarly, servers can buffer transactions that
require global coordination, executing an entire batch of these transactions
during a single round of global coordination.

>>>>

\section{Segmented Invariant Confluence}\seclabel{SegmentedInvariantConfluence}
\newcommand{\IsIclosed}{\textsc{IsIclosed}}

@@

\begin{example}\examplelabel{SegmentedZ2}
  Consider again the object $O = (\ints \times \ints, \join)$, transactions $T
  = \set{t_{x+1}, t_{y-1}}$, and invariant $I = \setst{(x, y)}{xy \leq 0}$ from
  \exampleref{Z2}, but now let the start state $s_0$ be $(-42, 42)$. $O$ is
  \emph{not} \sTIconfluent{} because the points $(0, 42)$ and $(42, 0)$ are
  reachable, and merging these points yields $(42, 42)$ which violates the
  invariant. However, $O$ is \sTISconfluent{} for $\Sigma = (I_1, T_1)$, $(I_2,
  T_2)$, $(I_3, T_3)$, $(I_4, T_4)$ where
  \begin{align*}
    I_1 &= \setst{(x, y)}{x < 0, y > 0}       &T_1 &= \set{t_{x+1}, t_{y-1}} \\
    I_2 &= \setst{(x, y)}{x \geq 0, y \leq 0} &T_2 &= \set{t_{x+1}, t_{y-1}} \\
    I_3 &= \setst{(x, y)}{x = 0}              &T_3 &= \set{t_{y-1}} \\
    I_4 &= \setst{(x, y)}{y = 0}              &T_4 &= \set{t_{x+1}}
  \end{align*}
  $\Sigma$ is illustrated in \figref{SegmentedZ2}. Clearly, $s_0$ satisfies the
  invariant, and $I_1, I_2, I_3, I_4$ cover $I$. Moreover, for every $(I_i,
  T_i) \in \Sigma$, we see that $O$ is \iclosed{I_i}, so $O$ is
  \sticonfluent{s}{T_i}{I_I} for every $s \in I_i$. Thus, $O$ is
  \sTISconfluent{}.
\end{example}

@@

\begin{itemize}
  \item
    If $t \notin T_i$, then $p_i$ initiates a round of global coordination to
    execute $t$. During global coordination, every server temporarily stops
    processing transactions and transitions to state $s = s_1 \join \ldots
    \join s_n$, the join of every server's state. Then, every server
    speculatively executes $t$ transitioning to state $t(s)$. If $t(s)$
    violates the invariant (i.e.  $\lnot I(t(s))$), then every server aborts
    $t$ and reverts to state $s$.  Then, $p_i$ replies to the client. If $t(s)$
    satisfies the invariant (i.e.  $I(t(s))$), then every server commits $t$
    and remains in state $t(s)$. Every server then chooses the first segment
    $(I_i, T_i) \in \Sigma$ such that $I_i(t(s))$ to be the new active segment.
    Note that such a segment is guaranteed to exist because the segment
    invariants cover $I$. Moreover, $\Sigma$ is ordered, so every server will
    deterministically pick the same active segment. In fact, an invariant of
    the system model is that at any given point of normal processing, every
    server has the same active segment.

  \item
    If $t \in T_i$, then $p_i$ executes $t$ immediately and without
    coordination.  If $t(s_i)$ satisfies the \emph{active} invariant (i.e.
    $I_i(t(s_i))$), then $p_i$ commits $t$, stays in state $t(s_i)$, and
    replies to the client. If $t(s_i)$ violates the \emph{global} invariant
    (i.e. $\lnot I(t(s_i))$), then $p_i$ aborts $t$, reverts to state $s_i$,
    and replies to the client. If $t(s_i)$ satisfies the global invariant but
    violates the active invariant (i.e.  $I(t(s_i))$ but $\lnot I_i(t(s_i))$),
    then $p_i$ reverts to state $s_i$ and initiates a round of global
    coordination to execute $t$, as described in the previous paragraph.
    Transaction execution is summarized in \algoref{TxnExecution}.
\end{itemize}

@@

\subsection{Interactive Decision Procedure}
In order for us to determine whether or not an object $O$ is \sTISconfluent{},
we have to determine whether or not $O$ is \invariantconfluent{} within each
segment $(I_i, T_i) \in \Sigma$. That is, we have determine if $O$ is
\sticonfluent{s}{T_i}{I_i} confluent for every state $s \in I_i$. Ideally, we
could leverage \algoref{InteractiveDecisionProcedure}, invoking it once per
segment. Unfortunately, \algoref{InteractiveDecisionProcedure} can only be used
to determine if $O$ is \sticonfluent{s}{T_i}{I_i} for a \emph{particular} state
$s \in I_i$, not for \emph{every} state $s \in I_i$. Thus, we would have to
invoke \algoref{InteractiveDecisionProcedure} $|I_i|$ times for every segment
$(I_i, T_i)$, which is clearly infeasible given that $I_i$ can be large or even
infinite.

@@

Using \thmref{ExtendedIclosureEquivalentIconfluence}, we develop
\algoref{ArbitraryStartInteractiveDecisionProcedure}: a natural generalization
of \algoref{InteractiveDecisionProcedure}.
\algoref{InteractiveDecisionProcedure} iteratively refines the set of
\emph{reachable} states whereas
\algoref{ArbitraryStartInteractiveDecisionProcedure} iteratively refines the
set of \emph{coreachable} states, but otherwise, the core of the two algorithms
is the same.\footnote{%
  Another small difference is that \IsIclosed{} behaves differently in
  \algoref{InteractiveDecisionProcedure} and
  \algoref{ArbitraryStartInteractiveDecisionProcedure}. In
  \algoref{ArbitraryStartInteractiveDecisionProcedure}, \IsIclosed{} returns a
  triple (closed, $s_1$, $s_2$). If closed is false, then $s_1, s_2 \in I$
  are two states not in $NR$ such that $I(s_1)$ and $I(s_2)$ but $\lnot I(s_1
  \join s_2)$. If no such states exist, then closed is true, and $s_1$ and
  $s_2$ are null.
}
Now, a segmented \invariantconfluence{} decision procedure, can simply invoke
\algoref{ArbitraryStartInteractiveDecisionProcedure} once on each segment.

@@

Third, we can integrate a couple of optimizations into our system model to
further reduce the amount of coordination it requires. First, if a server with
state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t \in I_i$
to execute, and $t(s_i)$ violates the active invariant but not the global
invariant, instead of initiating a round of global coordination, $p_i$ can
simply buffer $t$ for re-execution at a later time. While this increases the
latency required to execute $t$, it's possible that after other transactions
are executed, re-executing $t$ may lead to a state that either satisfies the
active invariant or violates the global invariant. In either case, a round of
global coordination is avoided. Similarly, servers can buffer transactions that
require global coordination, executing an entire batch of these transactions
during a single round of global coordination.
